Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-23 paper/paper.tex
added categories and moved Jarek to the right

\numberofauthors{3}
\author{
    \alignauthor Maciej Malawski\\
       \affaddr{AGH University of Science and Technology}\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{Krakow, Poland; and}\\
       \affaddr{University of Notre Dame}\\
       \affaddr{Center for Research Computing}\\
       \affaddr{111 ITC, Notre Dame, IN, USA}\\
       \email{mmalawsk@nd.edu}
    \alignauthor Jarek Nabrzyski\\
       \affaddr{University of Notre Dame}\\
       \affaddr{Center for Research Computing}\\
       \affaddr{111 ITC, Notre Dame, IN, USA}\\
       \email{naber@nd.edu}
    \alignauthor Gideon Juve and Ewa Deelman\\
       \affaddr{USC Information Sciences Institute}\\
       \affaddr{4676 Admiralty Way}\\
       \affaddr{Marina del Rey, CA, USA}\\
       \email{\{gideon,deelman\}@isi.edu}
}

@@

%\category{H.4}{Information Systems Applications}{Miscellaneous}
%\category{D.2.8}{Software Engineering}{Metrics}[complexity measures, performance measures]

%\terms{Theory}

\keywords{Scientific workflows, cloud computing, DAG scheduling, simulation}

@@

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
each workflow. The \textsc{PlanWorkflow} procedure first calls 
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
schedules tasks onto VMs, allocating new VMs when necessary. For each task 
in the workflow, the the least expensive slot is chosen to schedule the 
task so that it can be completed by  its deadline. VMs are allocated in 
blocks of one billing cycle (one hour) regardless of the size of the task. 
When computing the cost of scheduling a task on a given VM, the algorithm 
considers idle slots in blocks that were allocated for previous tasks to be 
free, while slots in new blocks cost the full price of a billing cycle. For 
example, if a task has a runtime of 10 minutes, and the price of a block is 
\$1, then the algorithm will either schedule the task on 
an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
or it will allocate a new block on an existing VM, or provision a new VM, for 
a cost of \$1. If the cost of slots on two different VMs is equal, then the 
slot with the earliest start time is chosen. To prevent too many VMs from 
being provisioned, the algorithm always prefers to extend the runtime of 
existing VMs before allocating new VMs. The result of this is that the 
algorithm will only allocate a new block if there are no idle slots on 
existing blocks large enough or early enough to complete the 
task by its deadline, and it will only allocate a new VM if it cannot add 
a block to an existing VM to complete the task by its deadline.

@@

%\begin{figure}[b!] 
%\centering
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-MONTAGE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-CYBERSHAKE-n-1000-8-dagh1-20m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-LIGO-n-1000-8-dagh1-40m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-GENOME-n-1000-8-dagh100-1500m0.pdf}
%\includegraphics[width=0.19\textwidth]{figures/pareto-score-SIPHT-n-1000-8-dagh5-50m0.pdf}
%\caption{Comparison of algorithms based on score, which  is computed as
%$\sum(1/(p+1))$ where priorities of subsequent workflows are $p=0,1,2,\ldots$
%(0 means highest priority).}
%\label{fig:score}
%\end{figure}

@@

Another lesson learned from our studies is the effect of task granularity on 
workflow ensembles. All the benchmark workflows that we tested have an average 
number of tasks (50--1000), but the Montage and CyberShake workflows have 
fine-grained tasks with execution times on the order of seconds to minutes, 
while LIGO, Epigenomics and SIPHT have coarser-grained tasks with runtimes 
ranging from several minutes to hours. We observed that the dynamic 
algorithms achieved better results on the fine grained workflows since the 
simple random scheduling strategy leads to good load-balancing and high 
resource utilization for these ensembles. On the other hand, when tasks are
longer, the dynamic algorithms tend to use up the budget by running tasks 
from low-priority workflows while the more sophisticated static scheduling 
strategy of SPSS is more circumspect. In the future we plan to investigate 
the effects of task granularity on ensemble scheduling in a more systematic
fashion.

>>>>

\numberofauthors{3}
\author{
    \alignauthor Maciej Malawski\\
       \affaddr{AGH University of Science and Technology}\\
       \affaddr{Dept. of Computer Science}\\
       \affaddr{Krakow, Poland; and}\\
       \affaddr{University of Notre Dame}\\
       \affaddr{Center for Research Computing}\\
       \affaddr{111 ITC, Notre Dame, IN, USA}\\
       \email{mmalawsk@nd.edu}
    \alignauthor Gideon Juve and Ewa Deelman\\
       \affaddr{USC Information Sciences Institute}\\
       \affaddr{4676 Admiralty Way}\\
       \affaddr{Marina del Rey, CA, USA}\\
       \email{\{gideon,deelman\}@isi.edu}
    \alignauthor Jarek Nabrzyski\\
       \affaddr{University of Notre Dame}\\
       \affaddr{Center for Research Computing}\\
       \affaddr{111 ITC, Notre Dame, IN, USA}\\
       \email{naber@nd.edu}
}

@@

\category{H.3.4}{Systems and Software}{Distributed systems}
\category{D.4.8}{Performance}{Simulation}

\terms{DAG, Scheduling, Simulation}

\keywords{Scientific workflows, Workflow ensembles,
Cloud computing, Infrastructure-as-a-Service}

@@

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for
each workflow. The \textsc{PlanWorkflow} procedure first calls 
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks according
to the procedure described above. Then, the \textsc{PlanWorkflow} procedure
schedules tasks onto VMs, allocating new VMs when necessary. For each task 
in the workflow, the least expensive slot is chosen to schedule the 
task so that it can be completed by  its deadline. VMs are allocated in 
blocks of one billing cycle (one hour) regardless of the size of the task. 
When computing the cost of scheduling a task on a given VM, the algorithm 
considers idle slots in blocks that were allocated for previous tasks to be 
free, while slots in new blocks cost the full price of a billing cycle. For 
example, if a task has a runtime of 10 minutes, and the price of a block is 
\$1, then the algorithm will either schedule the task on 
an existing VM that has an idle slot larger than 10 minutes for a cost of \$0, 
or it will allocate a new block on an existing VM, or provision a new VM, for 
a cost of \$1. If the cost of slots on two different VMs is equal, then the 
slot with the earliest start time is chosen. To prevent too many VMs from 
being provisioned, the algorithm always prefers to extend the runtime of 
existing VMs before allocating new VMs. The result of this is that the 
algorithm will only allocate a new block if there are no idle slots on 
existing blocks large enough or early enough to complete the 
task by its deadline, and it will only allocate a new VM if it cannot add 
a block to an existing VM to complete the task by its deadline.

@@

@@

Another lesson learned from our studies is the effect of task granularity on
workflow ensembles. The Montage and CyberShake workflows have fine-grained tasks
with execution times on the order of seconds to minutes, while LIGO, Epigenomics
and SIPHT have coarser-grained tasks with runtimes ranging from several minutes
to hours. We observed that the dynamic algorithms achieved better results on the
fine grained workflows since the simple random scheduling strategy leads to good
load-balancing and high resource utilization for these ensembles. On the other
hand, when tasks are longer, the dynamic algorithms tend to use up the budget by
running tasks from low-priority workflows while the more sophisticated static
scheduling strategy of SPSS is more circumspect. In the future we plan to
investigate the effects of task granularity on ensemble scheduling in a more
systematic fashion.
